---
title: "Forecast Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Project Setup

-   Install and Load Packages

```{r}

library(tidyverse)
library(lubridate)
library(tsibble)
library(feasts)
library(dplyr)
library(fpp3)
```

-   Import Dataset

```{r}
Oil <- read.csv("BrentOilPrices.csv")

```


## Data Wrangling

-   Aggregate time series to desired format for forecasting

```{r}

#Convert Date (chr) to (Date)
Oil$Date <- as.Date(Oil$Date, format = "%d-%b-%y")

```

-   Deal with missing data

```{r}
#No Missing Data but there is missing data between days
sum(is.na(Oil))
tail(Oil)

#fill gaps in Dates
Oil <- Oil %>%
  complete(Date = seq.Date(min(Date), max(Date), by="day")) %>%
  fill("Price")

Oil
```

-   Convert to tsibble

```{r}

Oil <- Oil %>%
  as_tsibble(index = Date)

Oil
```

-   Create new variables to aid in forecasting

```{r}

#Year Extraction
Oil <- Oil %>%
  mutate(Year = year(Date))

#Quarter
Oil <- Oil %>%
  mutate(Quarter = quarter(Date))

#Month
Oil <- Oil %>%
  mutate(Month = month(Date))

#Day of the week
Oil <- Oil %>%
  mutate(Weekday = wday(Date))

```

## Exploratory analysis and visualization for the dataset

```{r}
Oil %>% autoplot(Price)
```

```{r}
Oil %>% gg_season(Price, period = "year")
```

```{r}
Oil %>% gg_subseries(Price, period = "month")
Oil %>% gg_subseries(Price, period = "weekday")
Oil %>% gg_lag(Price, period = "weekday")
Oil %>% ACF(Price)
```

```{r}

ggplot(Oil)+
  geom_bar(aes(x=Month, y=Price, fill= Weekday), stat = "identity")

ggplot(Oil)+
  geom_bar(aes(x=Quarter, y=Price, fill= Month), stat = "identity")

ggplot(Oil)+
  geom_bar(aes(x=Weekday, y=Price, fill= Year), stat = "identity")
```

## Model fitting

- Split dataset into training and test sets 

```{r}

  #training/test split
  #15%-85% Split.

train <- Oil %>% filter(Date < '2015-03-28')
test <- Oil %>% filter(Date >= '2015-03-28')
```

- Fit Benchmark Models/Fit TSLM Model
  LM Fitting

```{r}

#train a few linear models and benchmark models
lm <- train %>% 
  model(
    lm = TSLM(Price ~ trend())
    )

glance(lm)


aug_lm <- augment(lm)
aug_lm


aug_lm %>% 
  ggplot(aes(x=Date)) +
  geom_line(aes(y = Price, color = "Data")) +
  geom_line(aes(y = .fitted, color = "Fitted")) +
  scale_color_manual(values = c(Data = "Black", Fitted = "Blue"))

lm %>% gg_tsresiduals()

aug_lm %>% features(.innov, unitroot_kpss)
#Not Stationary

train %>% features(Price, unitroot_kpss)
#Original Dat Not Stationary


#difference to convert to stationary, feature engineering (lags)
Oil_n <- Oil %>% 
  select(c(Price)) %>% 
  mutate(Diff_Price = difference(Price),
         P_lag1 = lag(Diff_Price, 1),
         P_lag2 = lag(Diff_Price, 2),
         P_lag3 = lag(Diff_Price, 3),
         P_lag4 = lag(Diff_Price, 4)) %>% 
  filter(!is.na(P_lag4))

Oil_n %>% autoplot(Diff_Price)

#high pvalue of kpss implies stationary
Oil_n %>% features(Diff_Price, unitroot_kpss)

train_lm <- Oil_n %>% filter(Date < '2015-03-28')
test_lm <- Oil_n %>% filter(Date >= '2015-03-28')

#train a few linear models and benchmark models
models <- train_lm %>% 
  model(
    lm_t_lag = TSLM(Diff_Price ~ trend() + P_lag1 + P_lag2 + P_lag3 + P_lag4),
    lm_lag = TSLM(Diff_Price ~ P_lag1 + P_lag2 + P_lag3 + P_lag4),
    mean = MEAN(Diff_Price),
    naive = NAIVE(Diff_Price),
    drift = RW(Diff_Price ~ drift()),
    lm = TSLM(Diff_Price ~ trend())
  )

glance(models)

TSLM <- train_lm %>%
  model(
    lm = TSLM(Diff_Price ~ trend())
  )
  

#view the regression summary
report(TSLM)

#inspect residuals
aug_TSLM <- augment(TSLM)

aug_TSLM %>% 
  ggplot(aes(x=Date)) +
  geom_line(aes(y = Diff_Price, color = "Data")) +
  geom_line(aes(y = .fitted, color = "Fitted")) +
  scale_color_manual(values = c(Data = "Black", Fitted = "Blue"))

#first step is to use gg_tsresiduals to assess residual autocorrelation, distribution and changing properties
TSLM %>% gg_tsresiduals()

aug_TSLM %>% features(.innov, unitroot_kpss)


#plot residuals against predictors to ensure we modeled correctly
train_lm %>% 
  left_join(residuals(TSLM), by = "Date") %>% 
  pivot_longer(c(P_lag1:P_lag4),
               names_to = "regressor", values_to = "x") %>% 
  ggplot(aes(x = x, y = .resid)) +
  geom_point() +
  facet_wrap(. ~ regressor, scales = "free_x") +
  labs(y = "Residuals", x = "")

#fitted values vs residuals looks like noise
aug_TSLM %>% 
  ggplot(aes(x = .fitted, y = .innov)) +
  geom_point() + labs(x = "Fitted", y = "Residuals")

#After Differencing Data Model AICc went from 59506.01 to -2603.548
#TSLM Is best model in Family
glance(lm)
glance(TSLM)
```

- Fit ETS Model/Comparing Model performance using accuracy measures  
  ETS Fitting

```{r}
#Fitting ETS Models

#fitting models
fit_ETS <- Oil %>% 
  model(SES = ETS(Price ~ error("A") + trend("N") + season("N")),
        Holt = ETS(Price ~ error("A") + trend("A") + season("N")),
        Damped = ETS(Price ~ error("A") + trend("Ad") + season("N")))

#review fitted models (AICc)
glance(fit_ETS)


#training/test split
#15%-85% Split.

train_ETS <- Oil %>% filter(Date < '2015-03-28')
test_ETS <- Oil %>% filter(Date >= '2015-03-28')

#use AICc to find me a model and also check out some other models
models <- train_ETS %>% 
  model(
    AAA = ETS(Price ~ error("A") + trend("A") + season("A")),
    AAM = ETS(Price ~ error("A") + trend("A") + season("M")),
    MAA = ETS(Price ~ error("M") + trend("A") + season("A")))

glance(models)

#MMA model performs best 
MAA <- train_ETS %>% 
  model(AAM = ETS(Price ~ error("A") + trend("A") + season("M")))

report(MAA)

MAA %>% gg_tsresiduals()

```

- Prophet Model

```{r}
#install.packages("fable.prophet")
library(fable.prophet)

fit <- train %>% 
  model(prophet(Oil),
        ARIMA(Oil),
        ETS(Oil))

fc <- fit %>% forecast(h = "3 years")
fc %>% accuracy(test)

fit_prophet <- train %>% 
  model(prophet(Oil))

fit_prophet %>% components() %>% autoplot()
```

- ARIMA Fitting

```{r}

#is this data stationary?
Oil %>% ACF(Price) %>% autoplot()
Oil %>% features(Price, unitroot_kpss) #low p-value implies non-stationary

#Data is not stationary

Oil %>% features(Price, unitroot_nsdiffs) # 2 differences suggested
#but dataset is not long enough for 2 seasonal differences

Oil %>% features(Price, list(unitroot_nsdiffs, feat_stl))

Oil %>% features(difference(Price,52), unitroot_kpss) #nope
#we will take 1 difference on seasonal based on length of dataset

s_diff_Oil <- Oil %>% mutate(s_diff_price = difference(Price, 52),
                                ns_s_diff_price = difference(s_diff_price))

#now data is stationary based on kpss test
s_diff_Oil %>% autoplot(ns_s_diff_price)
s_diff_Oil %>% ACF(ns_s_diff_price) %>% autoplot()
s_diff_Oil %>% features(ns_s_diff_price, unitroot_kpss)


train_arima <- s_diff_Oil %>% filter(Date < '2015-03-28')
test_arima <- s_diff_Oil %>% filter(Date >= '2015-03-28')

#try out a few ARIMA models and check for AICc
arima <- train_arima %>% 
  model(AR1 = ARIMA(Price ~ pdq(3,1,0) + PDQ(0,1,0)),
        AR2 = ARIMA(Price ~ pdq(4,1,0) + PDQ(0,1,0)),
        automatic = ARIMA(Price), #fable algorithm
        automatic_exhaustive = ARIMA(Price, stepwise = FALSE), #exhaustive search
        automatic_no_seas_exhaustive = ARIMA(Price ~ PDQ(0, 0, 0), stepwise = FALSE), 
        #exhaustive search no seasonal differences
        automatic_no_seas = ARIMA(Price ~ PDQ(0,0,0)), #fable algorithm no seasonal differencing
        MA = ARIMA(Price ~ pdq(p = 0, d = 1, q = 0:4))) #MA model

#compare aicc
glance(arima)

#review what models were developed
models <- arima 
View(arima)

#compare accuracy on best performing models
MA <- train_arima %>% 
  model(MA = ARIMA(Price ~ pdq(p = 0, d = 1, q = 0:4)))

glance(MA)

MA %>% gg_tsresiduals()
```
    
## Accuracy Measures:

```{r}

#Linear TSLM Best Model Performance against Test Data.

fc_TSLM <- TSLM %>% 
  forecast(new_data = test_lm)

fc_TSLM %>% autoplot(test_lm, level = NULL)

fc_TSLM %>% accuracy(Oil_n) # test accuracy

glance(TSLM)

#ETS MA Best Model Performance against Test Data.


fc_MAA <- MAA %>%
  forecast(new_data = test_ETS)

fc_MAA %>% autoplot(test_ETS, level = NULL)

fc_MAA %>% accuracy(Oil)

glance(MAA)


#ARIMA 

fc_MA <- MA %>%
  forecast(new_data = test_arima)

fc_MA %>% autoplot(test_arima, level = NULL)

fc_MA %>% accuracy(s_diff_Oil)

glance(MA)



#Selecting TSLM Model


```

```{r}

fc_MA <- MA %>% forecast(h = 400)
fc_MA %>% autoplot(Oil_n)
```
